---
title: "Styler tool unit testing"
author: "Daniel Contractor"
date: "2025-01-15"
output: html_document
---

##   **Unit Testing for Styler Tool**

#    Author: Daniel Contractor

#    Last updated:   15/01/2025

#    Purpose: This file contains unit tests for each of the individual functions the 'styler_functions.R'.
#    The goal of these tests is to demonstrate that each function performs as intended. 
#    They therfore also provide a useful starting point for debugging.

#    Order of Unit Tests:

#   1. Test for blankline_disassemble
#   2. Test for comment_disassemble
#   3. Test for capitalize_sql_keywords
#   4. Test for function_disassemble
#   5. Test for special_patterns_disassemble
#   6. Test for remove_whitespace
#   7. Test for insert_newlines_before_keywords_and_brackets 
#   8. Test for calculate_indentation
#   9. Test for restore_whitespace
#  10. Test for reassemble_special_patterns
#  11. Test for find_balanced_brackets
#  11. Test for reassemble_functions
#  12. Test for reassemble_comments
#  13. Test for remove_double_newlines
#  14. Test for restore_blank_lines
#  15. Test for standardize_sql
#  16. Test for capitalize_known_titles
#  17. Test for drop_table_if_exists



## Unit Tests

```{r}
source("cutdown_elegant_function_pairs.R")
```

#Test for blankline_disassemble
```{r}

sql_code <- "SELECT * FROM table;\n\nINSERT INTO table VALUES (1,2);\n\n"
expected_output <- "SELECT * FROM table;<><>b<><>\nINSERT INTO table VALUES (1,2);<><>b<><>\n"

result <- blankline_disassemble(sql_code)
stopifnot(result == expected_output)

```

#Test for comment_disassemble
```{r}
sql_code <- "SELECT * FROM table WHERE id = 1 -- This is a comment"

expected_output <- list(
  "SELECT * FROM table WHERE id = 1<><>c1<><>",
  c1 = " -- This is a comment"
)

result <- comment_disassemble(sql_code)

stopifnot (result[[1]] == expected_output[[1]] & result[['c1']] == expected_output[['c1']]) 

```

#Test for capitalize_sql_keywords
```{r}
sql_code <-list( "select col1, col2 from table where col3 = 5","(col1 + col2)")
result <- capitalize_sql_keywords(sql_code)[[1]]
expected <- "SELECT col1, col2 FROM TABLE WHERE col3 = 5"
stopifnot (result == expected) 
```

#Test for find_balanced_brackets

```{r}
# Test input
input_text <- "SELECT SUM(column1), AVG(column2), MAX(column3 + column4) FROM table WHERE COUNT(column5) > 10"

# Expected result: balanced brackets for SQL functions in the text
expected_result <- list(
  "(column1)", 
  "(column2)", 
  "(column3 + column4)", 
  "(column5)"
)

# Run the function
result <- find_balanced_brackets(input_text)

# Validate the result
stopifnot(length(result) == length(expected_result))  # Ensure the number of matches is correct
#stopifnot(all(result == expected_result))
for (i in seq_along(expected_result)) {
  stopifnot(result[[i]] == expected_result[[i]])  # Check each match individually
}# Ensure all matches are as expected

```

#Test for function_disassemble
```{r}
# Test for function_disassemble
sql_list <- list(
  "SELECT ABS(a), CONCAT(b, c) FROM table"
)

expected_output <- list(
  "SELECT ABS<><>f1<><>, CONCAT<><>f2<><> FROM table",
  f1 = "(a)",
  f2 = "(b, c)"
)

result <- function_disassemble(sql_list)

stopifnot(result[[1]] == expected_output[[1]] & result[['f1']] == expected_output[['f1']] & result[['f2']] == expected_output[['f2']] ) 
```


#Test for special_patterns_disassemble
```{r}

sql_text <- list("SELECT * FROM table WHERE DATA_COMPRESSION ON; CREATE NONCLUSTERED INDEX idx ON table (col1); CASE WHEN col1 = 'A' THEN 1 ELSE 0 END","c1"="comment 1", "c2=comment 2")

expected_output <- list("SELECT * FROM table WHERE DATA_COMPRESSION ON; {}{}y1{}{}; {}{}y2{}{}",
                        "y1" = "\nCREATE NONCLUSTERED INDEX idx ON table (col1)", "y2" = "CASE WHEN col1 = 'A' THEN 1 ELSE 0 END")

result <- special_patterns_disassemble(sql_text)
stopifnot(result[[1]] == expected_output[[1]])
stopifnot(result[[2]] == expected_output[[2]])
```


#Test for remove_whitespace
```{r}
sql_content <- c("  SELECT  *  \tFROM   table  WHERE  column1  =  value1  ", "Some comment")
expected_result <- c("SELECT * FROM table WHERE column1=value1", "Some comment")

result <- remove_whitespace(sql_content)
stopifnot (result[[1]] == expected_result[[1]] && result[[2]] == expected_result[[2]])
```

#Test for insert_newlines_before_keywords_and_brackets
```{r}
# Test for insert_newlines_before_keywords_and_brackets
sql_code <- list("SELECT * FROM table WHERE column1 = value1", "Some comment")
keywords_list <- c("SELECT", "FROM", "WHERE")
expected_result <- c("\nSELECT * \nFROM table \nWHERE column1 = value1", "Some comment")

result <- insert_newlines_before_keywords_and_brackets(sql_code, keywords_list)
stopifnot (result[[1]] == expected_result[[1]] && result[[2]] == expected_result[[2]]) 
```


#Test for calculate_indentation
```{r}

sql_code <- "WHERE EXISTS(
SELECT 1
FROM [IDI_Sandpit].[DL-MAA20XX-YY].[ref_diagnosis] AS r
)"

expected_output <- "WHERE EXISTS(\n    SELECT 1\n    FROM [IDI_Sandpit].[DL-MAA20XX-YY].[ref_diagnosis] AS r\n)"

result <- calculate_indentation(sql_code)
stopifnot(result == expected_output)
```

#Another example with multi-layer indentation
```{r}
sql_code <- "WITH setup (\nSELECT a\n, b \nFROM table\n) \nSELECT c\n, d\n, e \nFROM (\nSELECT c\n, f \nFROM setup \nWHERE EXISTS (\nSELECT z \nFROM table\n) \nGROUP BY x\n, y\n) AS subq \nWHERE condition \nAND (\ncondition2 \nOR condition3\n)"

expected_output <- "WITH setup (\n    SELECT a\n        , b\n    FROM table\n)\nSELECT c\n    , d\n    , e\nFROM (\n    SELECT c\n        , f\n    FROM setup\n    WHERE EXISTS (\n        SELECT z\n        FROM table\n    )\n    GROUP BY x\n        , y\n) AS subq\nWHERE condition\nAND (\n    condition2\n    OR condition3\n)"

result <- calculate_indentation(sql_code)
stopifnot(result == expected_output)
```



#Test for calculate_indentation
```{r}
input_string <- c("SELECT\n(user_id) FROM table \nWHERE column1 = value1")
expected_result <- c("SELECT\n(user_id) FROM table\nWHERE column1 = value1")

result <- calculate_indentation(input_string)
stopifnot (result[[1]] == expected_result[[1]])
```

#Test for restore_whitespace
```{r}
# Test for standard_nocapv2
sql_input <- "SELECT * FROM table WHERE id=1 AND value>=10 OR value<=5,order>100"
expected_output_standardized <- "SELECT * FROM table WHERE id = 1 AND value >= 10 OR value <= 5, order > 100"

result_standardized <- restore_whitespace(sql_input)

stopifnot (result_standardized == expected_output_standardized) 
```


#Test for reassemble_special_patterns
```{r}
sql_text <- list("SELECT * FROM table WHERE DATA_COMPRESSION ON; {}{}y11{}{}; {}{}y2{}{}",
                        "y11" = "\nCREATE NONCLUSTERED INDEX idx ON table (col1)", "y2" = "CASE WHEN col1 = 'A' THEN 1 ELSE 0 END")

expected_output <- list("SELECT * FROM table WHERE DATA_COMPRESSION ON; \nCREATE NONCLUSTERED INDEX idx ON table (col1); CASE WHEN col1 = 'A' THEN 1 ELSE 0 END")

result <- reassemble_special_patterns(sql_text)
stopifnot(result[[1]] == expected_output[[1]])
#stopifnot(result[[2]] == expected_output[[2]])
```

#Test for reassemble_functions
```{r}

sql_list <- list(
  "SELECT<><>f1<><>,<><>f2<><> FROM table",
  f1 = " ABS(a)",
  f2 = " CONCAT(b, c)"
)

expected_output  <- list(
  "SELECT ABS(a), CONCAT(b, c) FROM table"
)

result <- reassemble_functions(sql_list)

stopifnot (result[[1]] == expected_output[[1]] ) 
```

#Test for reassemble_comments
```{r}
disassembled <- list(
  "SELECT * FROM table WHERE id = <><>c1<><>",
  c1 = "This is a comment for the first placeholder"
)

expected_output <- "SELECT * FROM table WHERE id = This is a comment for the first placeholder"

result <- reassemble_comments(disassembled)

stopifnot (result == expected_output) 
```



# Test for remove_double_newlines
```{r}

sql_code <- "SELECT * FROM table;\n\nINSERT INTO table VALUES (1,2);\n\n"
expected_output <- "SELECT * FROM table;\nINSERT INTO table VALUES (1,2);\n"
result <- remove_double_newlines(sql_code)
stopifnot(result == expected_output)
```

# Test for restore_blank_lines
```{r}

sql_code <- "SELECT * FROM table;<><>b<><>INSERT INTO table VALUES (1,2);<><>b<><>"
expected_output <- "SELECT * FROM table;\nINSERT INTO table VALUES (1,2);\n"

result <- restore_blank_lines(sql_code)
stopifnot(result == expected_output)
```

# Test for standardize_sql
```{r}
sql_input <- " SELECT   *  FROM   table WHERE id  =  1  AND value   >=   10 OR value <=   5 ; "
expected_output_standardized <- " select * from table where id=1 and value>=10 or value<=5;"

result_standardized <- standardize_sql(sql_input)

stopifnot (result_standardized == expected_output_standardized) 
```

# Test for drop_table_if_exists
```{r}
sql_code <- "DROP TABLE my_table; DROP TABLE IF EXISTS other_table;"
result <- drop_table_if_exists(sql_code)
expected_output <-"DROP TABLE IF EXISTS my_table; DROP TABLE IF EXISTS other_table;"
stopifnot(result == expected_output)
```






